3D Magnetic modelling of ellipsoidal bodies

### 2017-02-21 manuscript/manuscript.tex
define the norm of the perturbed inverse Lambda matrix

Consider a magnetized ellipsoid immersed in
a uniform inducing magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm and
$D$ and $I$ are respectively, the declination and inclination of the 
local-geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This field represents the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents, 
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eqs. \ref{eq:A} and 
\ref{eq:coord_transformation}) of a point referred to the main 
coordinate system is defined as follows \citep{sharma1966, 
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of 
the magnetic scalar potential $V(\mathbf{r})$ given by:
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta} 
\mathbf{M}(\mathbf{r}^{\prime})^{\top} 
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:V-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc} 
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$, 
the integral is conducted over the variables 
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and 
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in $\unit{Am^{-1}}$).
Eq. \ref{eq:V-potential} is valid anywhere, 
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

@@

In the case of isotropic susceptibility, the resultant magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic}) may be determined by solving the
following linear system:
\begin{equation}
\mathbf{\Lambda}^{-1} \, \mathbf{M} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-K-isotropic-inverse-Lambda}
\end{equation}
where, according to Eq. \ref{eq:Lambda-K-isotropic},
\begin{equation}
\mathbf{\Lambda}^{-1} = \mathbf{V}
\left( \mathbf{I} - \chi \, \tilde{\mathbf{N}}^{\dagger} \right)
\mathbf{V}^{\top} \: .
\label{eq:inverse-Lambda}
\end{equation}

Consider a perturbed matrix $\delta \mathbf{\Lambda}^{-1}$ given by
\begin{equation}
\delta \mathbf{\Lambda}^{-1} = \mathbf{\Lambda}^{-1} - \mathbf{I}
\label{eq:perturbed-inverse-Lambda}
\end{equation}
and, similarly, a perturbed magnetization vector $\delta \mathbf{M}$
given by
\begin{equation}
\delta \mathbf{M} = \mathbf{M} - \breve{\mathbf{M}} \: ,
\label{eq:perturbed-M}
\end{equation}
where $\mathbf{M}$ is the true magnetization obtained from Eq. 
\ref{eq:M-K-isotropic-inverse-Lambda} and $\breve{\mathbf{M}}$
is the approximated magnetization obtained from Eq. \ref{eq:M-approx}.

>>>>

Consider a magnetized ellipsoid immersed in
a uniform inducing magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm (or 2-norm) and
$D$ and $I$ are respectively, the declination and inclination of the 
local-geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This field represents the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents, 
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eqs. \ref{eq:A} and 
\ref{eq:coord_transformation}) of a point referred to the main 
coordinate system is defined as follows \citep{sharma1966, 
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of 
the magnetic scalar potential $V(\mathbf{r})$ given by:
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta} 
\mathbf{M}(\mathbf{r}^{\prime})^{\top} 
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:V-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc} 
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$, 
the integral is conducted over the variables 
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and 
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in $\unit{Am^{-1}}$).
Eq. \ref{eq:V-potential} is valid anywhere, 
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

@@

In the case of isotropic susceptibility, the resultant magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic}) may be determined by solving the
following linear system:
\begin{equation}
\mathbf{\Lambda}^{-1} \, \mathbf{M} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-K-isotropic-inverse-Lambda}
\end{equation}
where, according to Eq. \ref{eq:Lambda-K-isotropic},
\begin{equation}
\mathbf{\Lambda}^{-1} = \mathbf{V}
\left( \mathbf{I} - \chi \, \tilde{\mathbf{N}}^{\dagger} \right)
\mathbf{V}^{\top} \: .
\label{eq:inverse-Lambda}
\end{equation}
As we have already pointed out, the approximated
magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
represents the particular case in which the matrix 
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic}),
and consequently the matrix $\mathbf{\Lambda}^{-1}$
(Eq. \ref{eq:inverse-Lambda}), are close to the identity.
Our goal in this subsection is to bound the norm of the
difference between the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda})
and the approximated magnetization $\breve{\mathbf{M}}$
(Eq. \ref{eq:M-approx}).

Consider a perturbed matrix $\delta \mathbf{\Lambda}^{-1}$ given by
\begin{equation}
\delta \mathbf{\Lambda}^{-1} = \mathbf{\Lambda}^{-1} - \mathbf{I}
\label{eq:perturbed-inverse-Lambda}
\end{equation}
and, similarly, a perturbed magnetization vector $\delta \mathbf{M}$
given by
\begin{equation}
\delta \mathbf{M} = \mathbf{M} - \breve{\mathbf{M}} \: .
\label{eq:perturbed-M}
\end{equation}
By using these two equations, we may rewrite that of the approximated
magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) as follows:
\begin{equation}
\left( \mathbf{\Lambda}^{-1} - \delta \mathbf{\Lambda}^{-1} \right)
\left( \mathbf{M} - \delta \mathbf{M} \right) = 
\chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: .
\label{eq:perturbed-system}
\end{equation}
Now, by subtracting the true magnetization 
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic-inverse-Lambda})
from this linear system (Eq. \ref{eq:perturbed-system})
and rearranging the terms, we obtain the following linear
system for the perturbed magnetization $\delta \mathbf{M}$
(Eq. \ref{eq:perturbed-M}):
\begin{equation}
\delta \mathbf{M} = - \delta \mathbf{\Lambda}^{-1} \mathbf{M} \: .
\label{eq:perturbed-M-system}
\end{equation}

By using the concept of vector norm and its corresponding
operator norm \citep{demmel1997, golub2013}, we may use Eq.
\ref{eq:perturbed-M-system} to write the following inequality:
\begin{equation}
\frac{\| \delta \mathbf{M} \|}{\| \mathbf{M} \|} \leq
\| \delta \mathbf{\Lambda}^{-1} \| \: .
\label{eq:perturbed-M-inequality}
\end{equation}
where $\| \delta \mathbf{M} \|$ and $\| \mathbf{M} \|$
denote Euclidean norms (or 2-norms) and the term 
$\| \delta \mathbf{\Lambda}^{-1} \|$ denotes the matrix 2-norm
of $\delta \mathbf{\Lambda}^{-1}$, which is given by
\begin{equation}
\| \delta \mathbf{\Lambda}^{-1} \| = \chi \,
\tilde{n}^{\dagger}_{33} \: ,
\label{eq:norm-perturbed-inverse-Lambda}
\end{equation}
where $\tilde{n}^{\dagger}_{33}$ is defined in the 
following sections, depending on the ellipsoid
type.
